---
title: "Homework 10"
author: "Emma Staut"
date: "4/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Part 1: Accident Data
Read in the "accident" data set.
```{r}
acc = read.csv("C:/Users/ecsta/Downloads/accident.csv")
head(acc)
```
1. What days of the week do accidents happen most often?
```{r}
#count how many accidents happen each day of the week
acc %>% group_by(DAY_WEEK) %>% tally()

#plot the number of accidents by day of the week
acc %>% ggplot(aes(x=DAY_WEEK)) + geom_bar()
```
As shown in the bar plot, accidents happen most often on the weekends. Saturday has the most accidents, followed by Friday and then Sunday.


2. What time of day do accidents happen most often?
```{r}
#check the values of HOUR
summary(acc$HOUR)

#since there are only 24 hours in a day, set HOUR <= 24
acc.hour = acc %>% filter(HOUR <= 24) %>%
  group_by(HOUR) %>% tally()

#plot the number of accidents by hour
acc %>% filter(HOUR <= 24) %>% ggplot(aes(x=HOUR)) + geom_bar()
```
As shown in the bar plot, the most accidents happen during the evening, with the most accidents around 6:00 pm. 

3. How many accidents have at least one drunk driver?
```{r}
#create a table with the number of drunk drivers per accident
table(acc$DRUNK_DR)
```
There are 8474 accidents with 1 drunk driver, 243 accidents with 2 drunk drivers, and 3 accidents with 3 drunk drivers. In total there are 8720 accidents with at least one drunk driver. 

# Part 2: Connecting Data
Read in the "person" data set. 
```{r}
per = read.csv("C:/Users/ecsta/Downloads/person.csv")

#create a subset using only drivers
per1 = per %>% filter(PER_TYP==1)
head(per1)
```

1. Join the accident and driver tables using 
```{r}

```

2. How many 